---
title: "Trump Endorsements"
output: github_document
---

Data Sources:

https://www.mcclatchydc.com/news/politics-government/election/midterms/article221145845.html

https://www.politico.com/story/2018/11/07/2018-election-results-trump-endorsements-967767

## Packages

```{r}
# devtools::install_github("favstats/tidytemplate")
# install.packages("pacman")
pacman::p_load(tidyverse, rvest, qdapRegex, ggthemes, tidytemplate, rtweet)
```

## Including Plots

```{r}
candidate <- read_html("https://www.mcclatchydc.com/news/politics-government/election/midterms/article221145845.html") %>% 
  html_nodes("#content-body- li") %>% 
  html_text()

mcclatch <- tibble(candidate) %>% 
    mutate(won = case_when(
        str_detect(candidate, "✓") ~ "Won", 
        str_detect(candidate, "✗") ~ "Lost", 
        T~ "Undecided")) %>% 
    mutate(type = case_when(
        str_detect(candidate, "House") ~ "House", 
        str_detect(candidate, "Senate|senator") ~ "Senate", 
        str_detect(candidate, "governor|gubernatorial") ~ "Governor")) %>% 
    mutate(incumbent = ifelse(str_detect(candidate, "incumbent|Incumbent"), "Incumbent", "Not Incumbent")) %>% 
    mutate(candidate = str_remove(candidate, "✓|✗") %>% str_trim()) %>% 
    mutate(candidate = str_extract(candidate, ".+?(?=,)")) %>% 
    mutate(candidate = str_remove(candidate, "\\(GOING TO RECOUNT\\) ")) %>% 
    mutate(candidate = str_remove(candidate, "Runoff: "))

mcclatch
```



```{r pressure, echo=FALSE}
politico_html <- read_html("https://www.politico.com/interactives/charts/chartwerk/zlncpz04.html?initialWidth=658&childId=chartwerk_zlncpz04&parentTitle=Trump%20said%20his%20name%20was%20on%20the%20ballot.%20Here%E2%80%99s%20how%20he%20fared.%20-%20POLITICO&parentUrl=https%3A%2F%2Fwww.politico.com%2Fstory%2F2018%2F11%2F07%2F2018-election-results-trump-endorsements-967767")
  
relevant_region <- politico_html %>% html_nodes("body") %>% 
  as.character() %>% 
  rm_between("%", "Trump report card", extract=TRUE) %>% 
  .[[1]] %>% 
  str_split("\\)") %>% 
  .[[1]]

candidate_names <- relevant_region %>% 
  str_split("Candidates") %>% 
  sapply(`[`,2) %>% 
  str_sub(5)

candidate_info <- relevant_region %>% 
  str_split("Candidates") %>% 
  sapply(`[`,1)  %>% 
  str_remove_all("[^[:alnum:] ]")

politico <- tibble(candidate_names, candidate_info) %>% 
  mutate(won = case_when(
        str_detect(candidate_info, "Won X") ~ "Won", 
        str_detect(candidate_info, "Lost X") ~ "Lost", 
        T ~ "Undecided")) %>% 
  mutate(type = case_when(
        str_detect(candidate_names, "House") ~ "House", 
        str_detect(candidate_names, "Senate|senator") ~ "Senate", 
        str_detect(candidate_names, "Gov") ~ "Governor")) %>% 
  mutate(candidate = str_extract(candidate_names, ".+?(?=\\()") %>% str_trim) %>% 
  select(candidate, everything(), -candidate_names, -candidate_info) %>% 
  drop_na()
  
politico
```

## Checking for Missings/Errors/Disagreements between Data Sources

Let's merge the datasets and see if we find any errors.

```{r}
politico %>% 
  full_join(mcclatch, by = "candidate") %>% 
  filter(is.na(won.x) | is.na(won.y))
```

Politico misspelled Patrick Morrisey.

Politico has more candidates but they did not include Mike Kelly.

Fix:

```{r}
politico <- politico %>% 
  mutate(candidate = ifelse(candidate == "Patrick Morrissey", "Patrick Morrisey", candidate))
```


Do the datasets disagree on whether the candidate ran for House or Senate?

```{r}
politico %>% 
  left_join(mcclatch, by = "candidate") %>% 
  filter(type.x != type.y)
```

Politico states that Martha McSally ran for the House of Representatives but this information seems to be wrong. She ran for Senate.

-> https://mcsallyforsenate.com/

Fix:

```{r}
politico <- politico %>% 
  mutate(type = ifelse(candidate ==  "Martha McSally", "Senate", type)) 
```


Do the datasets disagree on whether a race was lost/won?

```{r}
politico %>% 
  left_join(mcclatch, by = "candidate") %>% 
  filter(won.x != won.y)
```

Politico seems to not have updated Mike Dunleavy's win yet (via The Associated Press).

Data sources disagree on Diane Harkey. She is indeed projected to lose (via The Associated Press).

Rick Scott is in the lead in the Senate Race (via The Associated Press).

Data sources also disagree on Claudia Tenney. Probably lost the race (via The Associated Press).

Fix:

```{r}
politico <- politico %>% 
  mutate(won = ifelse(candidate == "Diane Harkey", "Lost", won)) %>% 
  mutate(won = ifelse(candidate == "Mike Dunleavy", "Won", won))  %>% 
  mutate(won = ifelse(candidate == "Rick Scott", "Won", won))  

mcclatch <- mcclatch %>% 
  mutate(won = ifelse(candidate == "Claudia Tenney", "Lost", won)) 
```

No errors left :)

We can now use the Politico dataset as base, as it is the most complete dataset:

```{r}
trump_endorsements <- politico  %>% 
  left_join(mcclatch, by = "candidate") %>% 
  rename(results = won.x) %>% 
  rename(type = type.x) %>% 
  select(candidate, results, type, incumbent) %>% 
  arrange(candidate)

trump_endorsements
```

## @GissiSim Data

@GissiSim on Twitter collected some Trump Endorsements as well. Let's check them out:

```{r, eval = F}
token <- readRDS(".rtweet.rds")

tml <- get_timelines("GissiSim", n = 500, token = token)



save_it(tml)
```


```{r}
tml <- load_it("data/tml.Rdata")

lastnames <- trump_endorsements %>% 
  mutate(lastname = str_remove(candidate, ".+?(?= )") %>% str_trim) %>% 
  .$lastname %>% paste0(collapse = "|")

tml %>% 
  filter(str_detect(text, "lost")) %>% 
  filter(!(str_detect(text, "Please let me know|That's 31|15 Trump|lost souls|democracy lost"))) %>%
  select(screen_name, text) %>% 
  mutate(lastnames = str_detect(text, lastnames)) %>% 
  filter(!(lastnames))
```


Karen Handel is not in the dataset but did receive an endorsement in 2017.

Randy is the first name of Randy Hultgren who is already in dataset.

Fix:


```{r}
trump_endorsements <- trump_endorsements %>% 
  rbind(c("Karen Handel", "Lost", "House", "Incumbent"))

save_it(trump_endorsements)
```

## Plotting the Data

Let's plot the data now :)

```{r, fig.width=10, fig.height=6}
gg_dat <- trump_endorsements %>% 
  group_by(results, type) %>% 
  tally() %>% 
  ungroup()%>% 
  group_by(type) %>% 
  mutate(total = sum(n)) %>% 
  ungroup() 

gg_text <- gg_dat %>% 
  filter(results == "Won") %>% 
  mutate(perc = tidytemplate::get_percentage(n, total, 2)) %>% 
  mutate(text = paste0(n, " out of ", total, " won (", perc,"%)"))

gg_dat %>% 
  ggplot(aes(results, n, fill = results)) +
  geom_col(alpha = 0.75) +
  facet_wrap(~type) +
  theme_hc() +
  scale_fill_colorblind() +
  guides(fill = F) +
  labs(x = "", y = "",
       caption = "Data: Politico & McClatchy Articles and Gissur Simonarson (@GissiSim)\nfavstats.eu; @favstats") +
  geom_text(aes(label = n), nudge_y = 1) +
  geom_text(data = gg_text, aes(x = 1.5, y = 31, label = text)) +
  ggtitle("How did Candidates endorsed by Trump fare in the 2018 Midterm Elections?\n")  +
  theme(legend.text = element_text(size = 8),
    axis.title = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0),
    plot.caption = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(size = 8), 
    ) 

ggsave_it(gg_freqs, width = 10, height = 6)
```



Plot by incumbents:

```{r, fig.width=10, fig.height=6}
gg_dat2 <- trump_endorsements %>% 
  group_by(results, incumbent) %>% 
  tally() %>% 
  ungroup()%>% 
  group_by(incumbent) %>% 
  mutate(total = sum(n)) %>% 
  ungroup() %>% 
  drop_na(incumbent)

gg_text2 <- gg_dat2 %>% 
  filter(results == "Won") %>% 
  mutate(perc = tidytemplate::get_percentage(n, total, 2)) %>% 
  mutate(text = paste0(n, " out of ", total, " won (", perc,"%)"))

gg_dat2 %>% 
  ggplot(aes(results, n, fill = results)) +
  geom_col(alpha = 0.75) +
  facet_wrap(~incumbent) +
  theme_hc() +
  scale_fill_colorblind() +
  guides(fill = F) +
  labs(x = "", y = "",
       caption = "Data: Politico & McClatchy Articles and Gissur Simonarson (@GissiSim)\nfavstats.eu; @favstats") +
  geom_text(aes(label = n), nudge_y = 1) +
  geom_text(data = gg_text2, aes(x = 1.5, y = 31, label = text)) +
  ggtitle("How did Candidates endorsed by Trump fare in the 2018 Midterm Elections?\n")  +
  theme(legend.text = element_text(size = 8),
    axis.title = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0),
    plot.caption = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(size = 8), 
    ) 

ggsave_it(gg_freqs2, width = 10, height = 6)
```



